---
title: "Chess"
author: "Jason Yoo"
date: "2/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(janitor)
library(tidyverse)
```
```{r}
lichess <- read.csv("data/games.csv") %>% clean_names()
internet_c <- read.csv("data//big_datasets/2016_CvC.csv") %>% clean_names()
internet_h <- read.csv("data/big_datasets/2016_CvH.csv") %>% clean_names()
internet <- rbind(internet_c, internet_h)
```

**About**\n

My name is Jason Yoo (Hwo Sun Yoo), and I am a junior in Leverett House studying Social Studies with a secondary in Statistics. In my later teenage years, I developed a bordering-unhealthy love of chess. Although I am not a great player, I have always been interested in studying the relationship between variables like opening moves, time spent, and differences in rating that might potentially affect the probability of winning a game of chess. 

For the project, I am using three different datasets. The first, named "lichess", is a public dataset from the popular online chess site, Lichess. It contains games between humans, regardless of playing level. The second and third are datasets containing information about computer versus computer or computer versus human games of varying playing levels. I have combined these two datasets into one, and named it "internet."

Through these datasets, I hope to study the significance of opening moves, color, moves spent per game, and rating differences between players on the probability of winning.